---
title: "Comparing Bayesian Models"
author: "Asma Becheikh"
date: ""
output: 
  html_document: 
    toc: yes
---

# Surgical: Institutional ranking
 
## Introduction 
    

I'm interesting in this Kernel in the comparison between two  models and checking which of them fits better  our   data. 

I'm considering a  data set of  mortality rates in 12 hospitals performing cardiac surgery in babies. 

```{r, prompt=T}
"dt" <-
list(n = c(47, 148, 119, 810,	211, 196, 148, 215, 207, 97, 256, 360),
        r  = c(0, 18, 8, 46, 8, 13, 9,	31, 14, 8, 29, 24),
        N = 12)
head(dt)
```



  
  # The Models 
  
  Since we have no prior information on the parameters we will use in the follwoing two bayesian models non-informative priors. 

Here's below the BUGS code of the them:  



+  Model1 

```{r, eval=F}
model
    {
    for( i in 1 : N ) {
		      p[i] ~ dbeta(1.0, 1.0)
			r[i] ~ dbin(p[i], n[i])
		   }
    
    }
```



+  Model2 

```{r, eval=F}
model
    {
    for( i in 1 : N ) {
				b[i] ~ dnorm(mu,tau)
				r[i] ~ dbin(p[i],n[i])
				logit(p[i]) <- b[i]
				}
			pop.mean <- exp(mu) / (1 + exp(mu))
			mu ~ dnorm(0.0,1.0E-6)
			sigma <- 1 / sqrt(tau)
			tau ~ dgamma(0.001,0.001)	   
    
    }
    
    ```
    

## `R` implementation

As mentioned above we will use for this analysis `R2OpenBUGS`, we need then to write these two models into two different `.txt` files. We need to write three `R` functions to initialize the Markov Chains. Here's then the `R` code that can be implemented. We will later run the `bugs` command and study the convergence of each generated Markov Chain.

Let's first load the needed packages


```{r, prompt=T}
library(R2OpenBUGS)
library(coda)
```

+ The initializing functions

```{r, prompt=T,eval=F}
inits1 <- function(){
  inits = list(p = c(0.1, 0.1, 0.1, 0.1,	0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1))
}

inits2 <- function(){
  inits = list(b = c(	0.1,	0.1,	0.1,	0.1,	0.1,	0.1,	0.1,	0.1,	0.1,	0.1,	0.1,	0.1), 
	     tau = 1, mu = 0)
}



```
+ The parameters 

```{r, prompt=T,eval=F}
params1<-"p"
params2<-c("b","tau","mu")



```
+ Creating the files of the BUGS code

```{r, prompt=F,eval=F}
###  Model1
sink('Mod1_cp.txt')
cat("
    model
    {
    for( i in 1 : N ) {
		      p[i] ~ dbeta(1.0, 1.0)
			r[i] ~ dbin(p[i], n[i])
		   }
    
    }
    ",fill=T)
sink()

###  Model2
sink('Mod2_cp.txt')
cat("
    model
    {
    for( i in 1 : N ) {
				b[i] ~ dnorm(mu,tau)
				r[i] ~ dbin(p[i],n[i])
				logit(p[i]) <- b[i]
				}
			pop.mean <- exp(mu) / (1 + exp(mu))
			mu ~ dnorm(0.0,1.0E-6)
			sigma <- 1 / sqrt(tau)
			tau ~ dgamma(0.001,0.001)	   
    
    }
    ",fill=T)
sink()

filename1<-"Mod1_cp.txt"
filename2<-"Mod2_cp.txt"


+ Running now the Markov Chains . We will run two MC with 1000 iterations with a burn-in equal to 900. 

```{r, eval=F}

out1<-bugs(dt,inits1,params1,filename1,codaPkg=F,
           n.thin =1, n.iter=10000,debug=F,
           n.chains = 3,working.directory=getwd(),
           OpenBUGS.pgm=OpenBUGS.pgm2, useWINE=T)
out2<-bugs(dt,inits2,params2,filename2,codaPkg=F,
           n.thin =1, n.iter=10000,debug=F,
           n.chains = 3,working.directory=getwd(),
           OpenBUGS.pgm=OpenBUGS.pgm, useWINE=T)


# Model Checking

```{r, echo=F}
load("C:/Users/asus/Documents/f10.RData")

```


## Model 1 (fixed effects)

We will then check the convergence of the Markov Chain contained in the `R` object  `out0`. 

We start then by the trace plot of the parameters

```{r, echo=T,prompt=T,collapse=F}
dim(out1$sims.array)
p=mcmc(out1$sims.array[,,1])
lambda=exp(p)
matplot(lambda,col=c("red","green","blue"),ylab=expression(lambda),type="l")

```
We proceed also by estimating the $widehat{R}$ and its confidence interval:

```
{r, prompt=T,collapse=F}
gelman.diag(list(p[,1],p[,2],p[,3]))
gelman.plot(list(p[,1],p[,2],p[,3]))
   

```
Let us proceed to the auto-correlation diagnostics

```{r, prompt=T,collapse=F}
autocorr(p)
autocorr.diag(p,0:10)
autocorr.plot(p)
```
We finally proceed to the Geweke diagnostics that show the stationary of the generated sample.

```{r, prompt=T,collapse=F}
geweke.diag(p)
```

We can then conclude that the Markov Chain in out1 converges to the posterior distribution.




## Model 2 (random effects)

We will then check the convergence of the Markov Chain contained in the `R` object `out2`. 

We start then by the trace plot of the parameters

```{r, prompt=T,collapse=T}
dim(out2$sims.array)
b=mcmc(out2$sims.array[,,1])
lambda1=exp(b)
matplot(lambda1,col=c("red","green","blue"),type="l")

tau=mcmc(out2$sims.array[,,2])
lambda2=exp(tau)
matplot(lambda2,col=c("red","green","blue"),type="l")


mu=mcmc(out2$sims.array[,,3])
matplot(mu,col=c("red","green","blue"),type="l")
```


We can't then conclude from these trace plots any abnormality in the convergence of the Markov Chain and let us then proceed to the others diagnostic methods. 

+ Estimation of $\widehat{R}$


```{r, prompt=T,collapse=T}
gelman.diag(list(b[,1],b[,2],b[,3]))
gelman.plot(list(b[,1],b[,2],b[,3])) 

gelman.diag(list(tau[,1],tau[,2],tau[,3]))
gelman.plot(list(tau[,1],tau[,2],tau[,3])) 

gelman.diag(list(mu[,1],mu[,2],mu[,3]))
gelman.plot(list(mu[,1],mu[,2],mu[,3])) 
```

We may need to run the MC with a number of iterations larger  than 10,000. We will try later 30,000 updates for the Markov Chain.

+ Auto-correlation procedure 


```{r, prompt=T,collapse=T}
autocorr.diag(b,0:10)
autocorr.plot(b)

autocorr.diag(tau,0:10)
autocorr.plot(tau)

autocorr.diag(mu,0:10)
autocorr.plot(mu)
```

We need then to increase the thin parameter when running the next  MC. 

+ a new update of the MC:

```{r,prompt=F,echo=T,eval=F}
out2a<-bugs(dt,inits2,params2,filename2,n.burnin = 25000,
           n.thin =5, n.iter=30000
           n.chains = 3
           )
```


 Let us then proceed with  another convergence diagnostics:


```{r, prompt=T,collapse=T}

b=mcmc(out2a$sims.array[,,1])
tau=mcmc(out2a$sims.array[,,2])
mu=mcmc(out2a$sims.array[,,3])

gelman.diag(list(b[,1],b[,2],b[,3]))
gelman.plot(list(b[,1],b[,2],b[,3])) 

gelman.diag(list(tau[,1],tau[,2],tau[,3]))
gelman.plot(list(tau[,1],tau[,2],tau[,3])) 

gelman.diag(list(mu[,1],mu[,2],mu[,3]))
gelman.plot(list(mu[,1],mu[,2],mu[,3])) 
```

```{r, prompt=T,collapse=T}
autocorr.diag(b,0:10)
autocorr.plot(b)

autocorr.diag(tau,0:10)
autocorr.plot(tau)

autocorr.diag(mu,0:10)
autocorr.plot(mu)
```


```{r, prompt=T,collapse=T}
geweke.diag(b)
geweke.diag(tau)
geweke.diag(mu)
```
The later MC convergences. We can considered it as a good estimate of the posterior distribution.

##Conclusion
 
 Finally when comparing the DIC of the two  models are
 
```{r, prompt=T,collapse=T}
out1$DIC
out2a$DIC
```


We can conclude that the model 1 ( fixed effects)  fits better the data. We conclude also that the mean of the number of deaths is following a Binomial  distribution . 
 
```{r, prompt=T,collapse=T}
out1

```













































